---
output: pdf_document
---

# FE590.  Assignment #4.


## Gang Ping Zhu
## `r format(Sys.time(), "%Y-%m-%d")`


# Instructions


When you have completed the assignment, knit the document into a PDF file, and upload _both_ the .pdf and .Rmd files to Canvas.

Note that you must have LaTeX installed in order to knit the equations below.  If you do not have it installed, simply delete the questions below.

# Question 1:
In this assignment, you will be required to find a set of data to run regression on.  This data set should be financial in nature, and of a type that will work with the models we have discussed this semester (hint: we didn't look at time series)  You may not use any of the data sets in the ISLR package that we have been looking at all semester.  Your data set that you choose should have both qualitative and quantitative variables. (or has variables that you can transform)

Provide a description of the data below, where you obtained it, what the variable names are and what it is describing.

# Answer 1:
The dataset is view of the finances of elementary and high schools in the United States. The information is from the United Sensus Bureau between 1993 and 2015. It was downloaded from kaggle.com (https://www.kaggle.com/noriuk/us-educational-finances/data) and the variables are as follows:

*STATE = The State which the revenue is provided 
*YEAR = The year in whichc the revenue is provided to the state
*TOTAL_REVENUE = TOTAL ELEMENTARY-SECONDARY REVENUE that was provided to the state
*FEDERAL_REVENUE = Total Revenue from Federal Sources
*STATE_REVENUE = Total Revenue from State Sources
*LOCAL_REVENUE = Total Revenue from Local Sources
*TOTAL_EXPENDITURE = TOTAL ELEMENTARY-SECONDARY EXPENDITURE. This value is calculated by adding the PROGRAM_CURRENT_EXPENDITURE, PROGRAM_NON_ELSEC_EXPENDITURE, CAPITAL_OUTLAY_EXPENDITURE, and payments/interest to the government. The last variable was not available in the data set so I've calculated it in the dataframe being used in our analysis. 
*PROGRAM_CURRENT_EXPENDITURE = TOTAL CURRENT SPENDING FOR ELEMENTARY-SECONDARY PROGRAMS. This value is calculated by adding the INSTRUCTION_EXPENDITURE, SUPPORT_SERVICES_EXPENDITURE, and PROGRAM_OTHER_EXPENDITURE.
*INSTRUCTION_EXPENDITURE = Total Current Spending for Instruction
*SUPPORT_SERVICES_EXPENDITURE = Total Current Spending for Support Services
*PROGRAM_OTHER_EXPENDITURE = TOTAL CURRENT SPENDING FOR OTHER ELEMENTARY-SECONDARY PROGRAMS
*PROGRAM_NON_ELSEC_EXPENDITURE = TOTAL CURRENT SPENDING FOR NONELEMENTARY-SECONDARY PROGRAMS
*CAPITAL_OUTLAY_EXPENDITURE = Total Capital Outlay Expenditure

# Question 2:
Pick a quantitative variable and fit at least four different models in order to predict that variable using the other predictors.  Determine which of the models is the best fit.  You will need to provide strong reason why the particular model you chose is the best one.  You will need to confirm the model you have selected provides the best fit and that you have obtained the best version of that particular model (i.e. subset selection or validation for example).  You need to convince the grader that you have chosen the best model.
```{r}
library(boot)
library(leaps)
library(glmnet)
library(caTools)
library(pls)
library(tree)
library(randomForest)
library(gbm)
library(gam)
library(splines)
library(MASS)
```

```{r}
df <- read.csv("elsec_summary.csv")
df
```
```{r}
df$PAYMENT_INT_TOTAL_EXP <- df$TOTAL_EXPENDITURE-df$PROGRAM_CURRENT_EXPENDITURE-df$PROGRAM_NON_ELSEC_EXPENDITURE-df$CAPITAL_OUTLAY_EXPENDITURE
df$PROG_PERC <- df$PROGRAM_CURRENT_EXPENDITURE/df$TOTAL_EXPENDITURE
df$NON__PROG_PERC <- df$PROGRAM_NON_ELSEC_EXPENDITURE/df$TOTAL_EXPENDITURE
df$CAP_OUT_PERC <- df$CAPITAL_OUTLAY_EXPENDITURE/df$TOTAL_EXPENDITURE
df$PAY_PERC <- df$PAYMENT_INT_TOTAL_EXP/df$TOTAL_EXPENDITURE
df$FEDERAL_PERC <- df$FEDERAL_REVENUE/df$TOTAL_REVENUE
df$STATE_PERC <- df$STATE_REVENUE/df$TOTAL_REVENUE
df$LOCAL_PERC <- df$LOCAL_REVENUE/df$TOTAL_REVENUE
df$HIGH_SOURCE <- ifelse(df$FEDERAL_REVENUE>df$STATE_REVENUE, "Federal", "State")
df$OVER_BUDGET <- ifelse(df$TOTAL_REVENUE>df$TOTAL_EXPENDITURE, "No", "Yes")
df
```
```{r}
#Running a model variable selection in order to determine which of the explanatory variables to use. I've removed state along with the Federal, State, and Local Revenue as the response variable that I'm interested in is the Total Revenue across the country. 

regexpfit.full=regsubsets(TOTAL_EXPENDITURE~YEAR+TOTAL_REVENUE+PROGRAM_CURRENT_EXPENDITURE+INSTRUCTION_EXPENDITURE+SUPPORT_SERVICES_EXPENDITURE+PROGRAM_OTHER_EXPENDITURE+PROGRAM_NON_ELSEC_EXPENDITURE+CAPITAL_OUTLAY_EXPENDITURE+PAYMENT_INT_TOTAL_EXP+PROG_PERC+NON__PROG_PERC+CAP_OUT_PERC+PAY_PERC+FEDERAL_PERC+STATE_PERC+LOCAL_PERC+HIGH_SOURCE+OVER_BUDGET,data=df, nvmax = 10, really.big = T)
t(summary(regexpfit.full)$which)
```
```{r}
plot(regexpfit.full)
plot(regexpfit.full, scale="Cp")
plot(regexpfit.full, scale="adjr2")
plot(regexpfit.full, scale="r2")
```

```{r}
#Running a model variable selection in order to determine which of the explanatory variables to use. I've removed state along with the Federal, State, and Local Revenue as the response variable that I'm interested in is the Total Revenue across the country. 

regfit.full=regsubsets(TOTAL_REVENUE~YEAR+TOTAL_EXPENDITURE+PROGRAM_CURRENT_EXPENDITURE+INSTRUCTION_EXPENDITURE+SUPPORT_SERVICES_EXPENDITURE+PROGRAM_OTHER_EXPENDITURE+PROGRAM_NON_ELSEC_EXPENDITURE+CAPITAL_OUTLAY_EXPENDITURE+PAYMENT_INT_TOTAL_EXP+PROG_PERC+NON__PROG_PERC+CAP_OUT_PERC+PAY_PERC+FEDERAL_PERC+STATE_PERC+LOCAL_PERC+HIGH_SOURCE+OVER_BUDGET,data=df, nvmax = 10, really.big = T)
t(summary(regfit.full)$which)
```
```{r}
plot(regfit.full)
plot(regfit.full, scale="Cp")
plot(regfit.full, scale="adjr2")
plot(regfit.full, scale="r2")
```

```{r}
reg.summary <- summary(regfit.full)
reg.summary
reg.summary$rsq
```

```{r}
clean.data <- df[,c("TOTAL_REVENUE", "YEAR", "INSTRUCTION_EXPENDITURE", "SUPPORT_SERVICES_EXPENDITURE", "CAPITAL_OUTLAY_EXPENDITURE", "PAYMENT_INT_TOTAL_EXP", "NON__PROG_PERC", "CAP_OUT_PERC", "FEDERAL_PERC", "LOCAL_PERC", "OVER_BUDGET")]
```

```{r}
set.seed(33)
glm.fit1 <- glm(TOTAL_REVENUE~
                  ns(YEAR,4)+
                  INSTRUCTION_EXPENDITURE+
                  SUPPORT_SERVICES_EXPENDITURE+
                  CAPITAL_OUTLAY_EXPENDITURE+
                  PAYMENT_INT_TOTAL_EXP+
                  NON__PROG_PERC+
                  CAP_OUT_PERC+
                  FEDERAL_PERC+
                  LOCAL_PERC+
                  OVER_BUDGET, data=df)
cv.err1 <- cv.glm(clean.data, glm.fit1)
cv.err1$delta
```

```{r}
glm.fit2 <- glm(TOTAL_REVENUE~
                  ns(YEAR,4)+
                  ns(TOTAL_EXPENDITURE,5)+
                  NON__PROG_PERC+
                  CAP_OUT_PERC+
                  FEDERAL_PERC+
                  LOCAL_PERC+
                  OVER_BUDGET,data=df)
cv.err2 <- cv.glm(df, glm.fit2)
cv.err2$delta
```

```{r}
glm.fit3 <- glm(TOTAL_REVENUE~
                  ns(YEAR,4)+
                  NON__PROG_PERC+
                  CAP_OUT_PERC+
                  FEDERAL_PERC+
                  LOCAL_PERC+
                  OVER_BUDGET,data=df)
cv.err3 <- cv.glm(df, glm.fit3)
cv.err3$delta
```
```{r}
glm.fit4 <- glm(TOTAL_REVENUE~
                  ns(YEAR,4)+
                  NON__PROG_PERC+
                  CAP_OUT_PERC+
                  FEDERAL_PERC+
                  LOCAL_PERC,data=df)
cv.err4 <- cv.glm(df, glm.fit4)
cv.err4$delta
```

```{r}
glm.fit <- glm(TOTAL_REVENUE~YEAR+TOTAL_EXPENDITURE,data=df)
gam.fit <- gam(TOTAL_REVENUE~s(YEAR,4)+s(TOTAL_EXPENDITURE,5)+STATE,data=df)
anova(glm.fit2, glm.fit, gam.fit)
```

```{r}
gam.m3=gam(wage~s(year,4)+s(age,5)+education,data=Wage)
par(mfrow=c(1,3))
plot(gam.m3,se=TRUE,col="blue")
plot.gam(gam1,se=TRUE,col="red")

gam.m1=gam(wage~s(age,5)+education,data=Wage)
gam.m2=gam(wage~year+s(age,5)+education,data=Wage)
anova(gam.m1,gam.m2,gam.m3,test="F")
```

```{r}
#Logistical Regression
logreg <- glm(PROG_PERC~TOTAL_REVENUE, family=binomial, data=df)
summary(logreg)
```

```{r}
#LDA.
logreg <- glm(Direction~Lag1+Lag2+Lag3+Lag4+Lag5+Volume, family=binomial, data=Weekly)
summary(logregweekly)
```

```{r}
#QDA.
logreg <- glm(Direction~Lag1+Lag2+Lag3+Lag4+Lag5+Volume, family=binomial, data=Weekly)
summary(logregweekly)
```

```{r}
#KNN.
logreg <- glm(Direction~Lag1+Lag2+Lag3+Lag4+Lag5+Volume, family=binomial, data=Weekly)
summary(logregweekly)
```

#Question 3:

Do the same approach as in question 2, but this time for a qualitative variable.
```{r}
#Logistical Regression
logreg <- glm(TOTAL_REVENUE~TOTAL_EXPENDITURE, family=binomial, data=df)
summary(logreg)
```

```{r}
#LDA.
logreg <- glm(Direction~Lag1+Lag2+Lag3+Lag4+Lag5+Volume, family=binomial, data=Weekly)
summary(logregweekly)
```

```{r}
#QDA.
logreg <- glm(Direction~Lag1+Lag2+Lag3+Lag4+Lag5+Volume, family=binomial, data=Weekly)
summary(logregweekly)
```

```{r}
#KNN.
logreg <- glm(Direction~Lag1+Lag2+Lag3+Lag4+Lag5+Volume, family=binomial, data=Weekly)
summary(logregweekly)
```

#Question 4:

(Based on ISLR Chapter 9 #7) In this problem, you will use support vector approaches in order to
predict whether a given car gets high or low gas mileage based on the
Auto
data set.

##(a)
Create a binary variable that takes on a 1 for cars with gas mileage above the median, and a 0 for cars with gas mileage below the median.
```{r}
library(ISLR)
gas.med = median(Auto$mpg)
new.var = ifelse(Auto$mpg > gas.med, 1, 0)
Auto$mpglevel = as.factor(new.var)
```

##(b)
Fit a support vector classifier to the data with various values of cost, in order to predict whether a car gets high or low gas mileage. Report the cross-validation errors associated with different values of this parameter. Comment on your results.
```{r}
library(e1071)

set.seed(999)
tune.out = tune(svm, mpglevel ~ ., data = Auto, kernel = "linear", ranges = list(cost = c(0.01, 
    0.1, 1, 5, 10, 100)))
summary(tune.out)
```
##(c)
Now repeat for (b), this time using SVMs with radial and polynomial basis kernels, with different values of gamma and degree and cost. Comment on your results.

```{r}

set.seed(333)
tune.out = tune(svm, mpglevel ~ ., data = Auto, kernel = "polynomial", ranges = list(cost = c(0.1, 
    1, 5, 10), degree = c(2, 3, 4)))
summary(tune.out)

set.seed(463)
tune.out = tune(svm, mpglevel ~ ., data = Auto, kernel = "radial", ranges = list(cost = c(0.1, 
    1, 5, 10), gamma = c(0.01, 0.1, 1, 5, 10, 100)))
summary(tune.out)
```

##(d)
Make some plots to back up your assertions in (b) and (c). Hint: In the lab, we used the plot() function for svm objects only in cases with p=2 When p>2,you can use the plot() function to create plots displaying pairs of variables at a time. Essentially, instead of typing plot(svmfit , dat) where svmfit contains your fitted model and dat is a data frame containing your data, you can type plot(svmfit , dat, x1???x4) in order to plot just the first and fourth variables. However, you must replace x1 and x4 with the correct variable names. To find out more, type ?plot.svm.

```{r}
svm.linear = svm(mpglevel ~ ., data = Auto, kernel = "linear", cost = 1)
svm.poly = svm(mpglevel ~ ., data = Auto, kernel = "polynomial", cost = 10, 
    degree = 2)
svm.radial = svm(mpglevel ~ ., data = Auto, kernel = "radial", cost = 10, gamma = 0.01)
plotpairs = function(fit) {
    for (name in names(Auto)[!(names(Auto) %in% c("mpg", "mpglevel", "name"))]) {
        plot(fit, Auto, as.formula(paste("mpg~", name, sep = "")))
    }
}
plotpairs(svm.linear)

plotpairs(svm.poly)
plotpairs(svm.radial)
```

