---
output: pdf_document
---

# FE590.  Assignment #4.


## Gang Ping Zhu
## `r format(Sys.time(), "%Y-%m-%d")`


# Instructions


When you have completed the assignment, knit the document into a PDF file, and upload _both_ the .pdf and .Rmd files to Canvas.

Note that you must have LaTeX installed in order to knit the equations below.  If you do not have it installed, simply delete the questions below.

# Question 1:
In this assignment, you will be required to find a set of data to run regression on.  This data set should be financial in nature, and of a type that will work with the models we have discussed this semester (hint: we didn't look at time series)  You may not use any of the data sets in the ISLR package that we have been looking at all semester.  Your data set that you choose should have both qualitative and quantitative variables. (or has variables that you can transform)

Provide a description of the data below, where you obtained it, what the variable names are and what it is describing.

# Answer 1:
The dataset is view of the interest rates across from the Federal Reserve of the United States. The information provides daily interest rate information for a variety of core interest rates, such as T-bills and the federal funds rate. For some rates, this dataset extends all the way back to 1954 but as we're looking at the dataset overall, some of those rates will not be used as it's an incomplete view across the data set. It was downloaded from the Federal Reserve (https://www.federalreserve.gov/releases/h15/) and the variables are as follows:

*time_period - This is the date of the recorded data points
*federal_funds - Federal funds, often referred to as fed funds, are excess reserves that commercial banks and other financial institutions deposit at regional Federal Reserve banks; these funds can be lent, then, to other market participants with insufficient cash on hand to meet their lending and reserve needs.
*financial and nonfinancial commercial paper - Commercial paper is an unsecured, short-term debt instrument issued by a corporation, typically for the financing of accounts receivable, inventories and meeting short-term liabilities.
*prime_rate - The prime rate is the interest rate that commercial banks charge their most credit-worthy customers. Generally, a bank's best customers consist of large corporations. The prime interest rate, or prime lending rate, is largely determined by the federal funds rate, which is the overnight rate that banks use to lend to one another; the prime rate is also important for individual borrowers, as the prime rate directly affects the lending rates available for a mortgage, small business loan or personal loan.
*discount_rate - The discount rate is the interest rate charged to commercial banks and other depository institutions for loans received from the Federal Reserve's discount window. The discount rate also refers to the interest rate used in discounted cash flow analysis to determine the present value of future cash flows.
*treasury_bill - A Treasury bill (T-Bill) is a short-term debt obligation backed by the Treasury Dept. of the U.S. government with a maturity of less than one year, sold in denominations of $1,000 up to a maximum purchase of $5 million. T-bills have various maturities and are issued at a discount from par.
*treasury_constant_maturity - Constant maturity is an adjustment for equivalent maturity, used by the Federal Reserve Board to compute an index based on the average yield of various Treasury securities maturing at different periods. Constant maturity yields on Treasuries are obtained by the U.S. Treasury on a daily basis through interpolation of the Treasury yield curve, which in turn is based on closing bid-yields of actively-traded Treasury securities.
*inflation_indexed_treasury_constant_maturity - An adjustment for equivalent maturity, used by the Federal Reserve Board to compute an index based on the average inflation.
*inflation_indexed_long_term_average - This is the average value of the inflation over time. 

# Question 2:
Pick a quantitative variable and fit at least four different models in order to predict that variable using the other predictors.  Determine which of the models is the best fit.  You will need to provide strong reason why the particular model you chose is the best one.  You will need to confirm the model you have selected provides the best fit and that you have obtained the best version of that particular model (i.e. subset selection or validation for example).  You need to convince the grader that you have chosen the best model.
```{r}
library(boot)
library(leaps)
library(glmnet)
library(caTools)
library(pls)
library(tree)
library(randomForest)
library(gbm)
library(gam)
library(splines)
library(MASS)
library(class)
```

```{r}
setwd("C:/Users/gang.ping.m.zhu/OneDrive - Accenture/Stevens/FE 590/A4")
int.rate.df <- read.csv("data.csv", header=TRUE, na.strings=-9999)
names(int.rate.df)
int.rate.clean.df <- int.rate.df[, c("month_1_nonfinancial_commercial_paper",
                                     "month_2_nonfinancial_commercial_paper", 
                                     "month_3_nonfinancial_commercial_paper",
                                     "month_1_financial_commercial_paper",                  
                                     "month_2_financial_commercial_paper",
                                     "month_3_financial_commercial_paper", 
                                     "prime_rate",
                                     "week_4_treasury_bill",
                                     "month_3_treasury_bill",
                                     "month_6_treasury_bill",
                                     "year_1_treasury_bill",
                                     "month_1_treasury_constant_maturity" ,
                                     "month_3_treasury_constant_maturity",
                                     "month_6_treasury_constant_maturity" ,
                                     "year_1_treasury_constant_maturity",
                                     "year_2_treasury_constant_maturity",
                                     "year_3_treasury_constant_maturity",
                                     "year_5_treasury_constant_maturity",
                                     "year_7_treasury_constant_maturity",
                                     "year_10_treasury_constant_maturity",
                                     "year_20_treasury_constant_maturity",
                                     "year_30_treasury_constant_maturity",
                                     "year_5_inflation_indexed_treasury_constant_maturity", 
                                     "year_7_inflation_indexed_treasury_constant_maturity",
                                     "year_10_inflation_indexed_treasury_constant_maturity",
                                     "year_20_inflation_indexed_treasury_constant_maturity", 
                                     "year_30_inflation_indexed_treasury_constant_maturity",
                                     "inflation_indexed_long_term_average")]
int.rate.clean.df <- na.omit(int.rate.clean.df)
head(int.rate.clean.df)
```
```{r}
#Running a model variable selection in order to determine which of the explanatory variables to use. I've removed a few variables that are almost linear. The response variable that I'm interested in is the prime rate. 
regfit.full <- regsubsets(prime_rate~.,data=int.rate.clean.df, nvmax = 15, really.big = T)
t(summary(regfit.full)$which)

```
```{r}
reg.summary <- summary(regfit.full)
reg.summary
reg.summary$rsq
```
```{r}
plot(regfit.full)
plot(regfit.full, scale="Cp")
plot(regfit.full, scale="adjr2")
plot(regfit.full, scale="r2")
```
```{r}
#Model 1 for Prime Rate
set.seed(33)
glm.fit1 <- glm(prime_rate~
                  month_2_nonfinancial_commercial_paper+
                  month_1_financial_commercial_paper+
                  month_6_treasury_bill+
                  year_3_treasury_constant_maturity+
                  year_5_treasury_constant_maturity+
                  year_7_treasury_constant_maturity+
                  year_10_treasury_constant_maturity+
                  year_20_treasury_constant_maturity+
                  year_30_treasury_constant_maturity+
                  year_5_inflation_indexed_treasury_constant_maturity+
                  year_7_inflation_indexed_treasury_constant_maturity+
                  year_10_inflation_indexed_treasury_constant_maturity+
                  year_20_inflation_indexed_treasury_constant_maturity+
                  inflation_indexed_long_term_average, data=int.rate.clean.df)
cv.err1 <- cv.glm(int.rate.clean.df, glm.fit1)
cv.err1$delta
# 0.001241602 0.001241588
```

```{r}
#Model 2 for Prime Rate
glm.fit2 <- glm(prime_rate~
                  month_1_nonfinancial_commercial_paper+
                  year_5_treasury_constant_maturity+
                  year_7_treasury_constant_maturity+
                  year_10_treasury_constant_maturity+
                  year_20_treasury_constant_maturity+
                  year_30_treasury_constant_maturity+
                  year_5_inflation_indexed_treasury_constant_maturity+
                  year_7_inflation_indexed_treasury_constant_maturity+
                  year_10_inflation_indexed_treasury_constant_maturity+
                  year_20_inflation_indexed_treasury_constant_maturity+
                  inflation_indexed_long_term_average, data=int.rate.clean.df)
cv.err2 <- cv.glm(int.rate.clean.df, glm.fit2)
cv.err2$delta
# 0.001690060 0.001690047
```

```{r}
# Model 3 for Prime Rate
glm.fit3 <- glm(prime_rate~
                  month_2_nonfinancial_commercial_paper+
                  month_1_financial_commercial_paper+
                  month_6_treasury_bill+
                  year_3_treasury_constant_maturity+
                  year_5_treasury_constant_maturity+
                  year_7_treasury_constant_maturity+
                  year_10_treasury_constant_maturity+
                  year_20_treasury_constant_maturity+
                  year_30_treasury_constant_maturity+
                  year_5_inflation_indexed_treasury_constant_maturity+
                  year_7_inflation_indexed_treasury_constant_maturity+
                  year_10_inflation_indexed_treasury_constant_maturity+
                  year_20_inflation_indexed_treasury_constant_maturity+
                  inflation_indexed_long_term_average+
                  poly(inflation_indexed_long_term_average, 2),data=int.rate.clean.df)
cv.err3 <- cv.glm(int.rate.clean.df, glm.fit3)
cv.err3$delta
#0.001186794 0.001186780
```

```{r}
# Model 4 for Prime Rate
glm.fit4<- glm(prime_rate~
                  month_2_nonfinancial_commercial_paper+
                  month_1_financial_commercial_paper+
                  month_6_treasury_bill+
                  year_3_treasury_constant_maturity+
                  year_5_treasury_constant_maturity+
                  year_7_treasury_constant_maturity+
                  year_10_treasury_constant_maturity+
                  year_20_treasury_constant_maturity+
                  year_30_treasury_constant_maturity+
                  year_5_inflation_indexed_treasury_constant_maturity+
                  year_7_inflation_indexed_treasury_constant_maturity+
                  year_10_inflation_indexed_treasury_constant_maturity+
                  year_20_inflation_indexed_treasury_constant_maturity+
                  inflation_indexed_long_term_average+
                  poly(inflation_indexed_long_term_average, 2)+
                  poly(year_20_inflation_indexed_treasury_constant_maturity, 2),data=int.rate.clean.df)
cv.err4 <- cv.glm(int.rate.clean.df, glm.fit4)
cv.err4$delta
#0.001153627 0.001153613
```

```{r}
anova(glm.fit1,glm.fit2,glm.fit3,glm.fit4)
```

Based on the computed LOOCV errors, model 4 is the best. 

#Question 3:

Do the same approach as in question 2, but this time for a qualitative variable.

```{r}
# Transforming the information into a quanitive variable
new.col.transform <- data.frame(int.rate.clean.df$inflation_indexed_long_term_average)
new.col.transform <- data.frame(new.col.transform[-1,])
new.col.transform <- rbind(new.col.transform, c(0))
int.rate.clean.df$inflation_indexed_long_term_average_previous <- new.col.transform
int.rate.clean.df$direction <- ifelse(int.rate.clean.df$inflation_indexed_long_term_average>=int.rate.clean.df$inflation_indexed_long_term_average_previous, "Up", "Down")
# int.rate.clean.df$inflation_indexed_long_term_average_previous <- NULL
```

```{r}
# Dividing into a training set and validation set
n = 1:floor(nrow(int.rate.clean.df)/2)
TrainingSet = int.rate.clean.df[n, ]
ValidationSet = int.rate.clean.df[ - n, ]
dim(TrainingSet)
names(TrainingSet)
dim(ValidationSet)
names(ValidationSet)
```

```{r}
# Model 1 for Infation Indexed Long Term Average Direction (Quantative Variable)
# LDA for Inflaction Indexed Long Term Average Direction
inf.d.Val = ValidationSet$direction

inf.d.lda.fit1 <- lda(direction~
                  month_2_nonfinancial_commercial_paper+
                  month_1_financial_commercial_paper+
                  month_6_treasury_bill+
                  prime_rate+
                  year_3_treasury_constant_maturity+
                  year_5_treasury_constant_maturity+
                  year_7_treasury_constant_maturity+
                  year_10_treasury_constant_maturity+
                  year_20_treasury_constant_maturity+
                  year_30_treasury_constant_maturity+
                  year_5_inflation_indexed_treasury_constant_maturity+
                  year_7_inflation_indexed_treasury_constant_maturity+
                  year_10_inflation_indexed_treasury_constant_maturity+
                  year_20_inflation_indexed_treasury_constant_maturity+
                  inflation_indexed_long_term_average, data=TrainingSet)

inf.d.lda.pred=predict(inf.d.lda.fit1,ValidationSet)
inf.d.lda.class=inf.d.lda.pred$class
table(inf.d.lda.class,inf.d.Val)
inf.d.LDA = round(mean(inf.d.lda.class==inf.d.Val)*100, 4)
inf.d.LDA
#Success rate of 49.6088%

#QDA for Inflaction Indexed Long Term Average Direction
inf.d.qda.fit1 <- qda(direction~
                        month_2_nonfinancial_commercial_paper+
                        month_1_financial_commercial_paper+
                        month_6_treasury_bill+
                        prime_rate+
                        year_3_treasury_constant_maturity+
                        year_5_treasury_constant_maturity+
                        year_7_treasury_constant_maturity+
                        year_10_treasury_constant_maturity+
                        year_20_treasury_constant_maturity+
                        year_30_treasury_constant_maturity+
                        year_5_inflation_indexed_treasury_constant_maturity+
                        year_7_inflation_indexed_treasury_constant_maturity+
                        year_10_inflation_indexed_treasury_constant_maturity+
                        year_20_inflation_indexed_treasury_constant_maturity+
                        inflation_indexed_long_term_average, data=TrainingSet)

inf.d.qda.pred=predict(inf.d.qda.fit1,ValidationSet)
inf.d.qda.class=inf.d.qda.pred$class
table(inf.d.qda.class,inf.d.Val)
inf.d.QDA = round(mean(inf.d.qda.class==inf.d.Val)*100, 4)
inf.d.QDA
#Success rate of 52.7387%

#KNN for Inflaction Indexed Long Term Average Direction

var1 <- c("month_2_nonfinancial_commercial_paper",
         "month_1_financial_commercial_paper",
         "month_6_treasury_bill",
         "prime_rate",
         "year_3_treasury_constant_maturity",
         "year_5_treasury_constant_maturity",
         "year_7_treasury_constant_maturity",
         "year_10_treasury_constant_maturity",
         "year_20_treasury_constant_maturity",
         "year_30_treasury_constant_maturity",
         "year_5_inflation_indexed_treasury_constant_maturity",
         "year_7_inflation_indexed_treasury_constant_maturity",
         "year_10_inflation_indexed_treasury_constant_maturity",
         "year_20_inflation_indexed_treasury_constant_maturity",
         "inflation_indexed_long_term_average")
var2 <- c("direction")
train.var1 <- TrainingSet[var1]
test.var1 <- ValidationSet[var1]
test.var1 <- test.var1[var1]
train.dep1 <- TrainingSet[var2]
test.dep1 <- ValidationSet[var2]

KNN.Multi <- rep(NA,50)
for (i in 1:50){
  set.seed(1)
  inf.knn.pred <- knn(train.var1,test.var1,train.dep1$direction,k = i)
  KNN.Multi[i] <- mean(inf.knn.pred==test.dep1$direction)
}
KN <- c(1:50)
KNN.Multi.KN <- cbind(KNN.Multi,KN)
inf.knn <- KNN.Multi.KN[which.max(KNN.Multi), ]
inf.knn
inf.knn.NoLag <- round(mean(inf.knn[1])*100, 4)
inf.knn.NoLag

#Success rate of 51.1737% when K = 6
```

```{r}
# Model 2 for Infation Indexed Long Term Average Direction
# LDA for Inflaction Indexed Long Term Average Direction

inf.d.lda.fit2 <- lda(direction~
                        month_1_nonfinancial_commercial_paper+
                        month_2_nonfinancial_commercial_paper+
                        month_2_financial_commercial_paper+
                        month_3_financial_commercial_paper+
                        prime_rate+
                        week_4_treasury_bill+
                        month_3_treasury_bill+
                        month_1_treasury_constant_maturity+
                        month_3_treasury_constant_maturity+
                        month_6_treasury_constant_maturity+
                        year_30_inflation_indexed_treasury_constant_maturity+
                        inflation_indexed_long_term_average, data=TrainingSet)

inf.d.lda.pred2 <- predict(inf.d.lda.fit2,ValidationSet)
inf.d.lda.class2 <- inf.d.lda.pred2$class
table(inf.d.lda.class2,inf.d.Val)
inf.d.LDA2 = round(mean(inf.d.lda.class2==inf.d.Val)*100, 4)
inf.d.LDA2
#Success rate of 48.5133%

#QDA for Inflaction Indexed Long Term Average Direction
inf.d.qda.fit2 <- qda(direction~
                        month_1_nonfinancial_commercial_paper+
                        month_2_nonfinancial_commercial_paper+
                        month_2_financial_commercial_paper+
                        month_3_financial_commercial_paper+
                        prime_rate+
                        week_4_treasury_bill+
                        month_3_treasury_bill+
                        month_1_treasury_constant_maturity+
                        month_3_treasury_constant_maturity+
                        month_6_treasury_constant_maturity+
                        year_30_inflation_indexed_treasury_constant_maturity+
                        inflation_indexed_long_term_average, data=TrainingSet)

inf.d.qda.pred2 <- predict(inf.d.qda.fit2,ValidationSet)
inf.d.qda.class2 <- inf.d.qda.pred2$class
table(inf.d.qda.class2,inf.d.Val)
inf.d.QDA2 = round(mean(inf.d.qda.class2==inf.d.Val)*100, 4)
inf.d.QDA2
#Success rate of 51.9562%

#KNN for Inflaction Indexed Long Term Average Direction

var1.2 <- c("month_1_nonfinancial_commercial_paper",
         "month_2_nonfinancial_commercial_paper",
         "month_2_financial_commercial_paper",
         "month_3_financial_commercial_paper",
         "prime_rate",
         "week_4_treasury_bill",
         "month_3_treasury_bill",
         "month_1_treasury_constant_maturity",
         "month_3_treasury_constant_maturity",
         "month_6_treasury_constant_maturity",
         "year_30_inflation_indexed_treasury_constant_maturity",
         "inflation_indexed_long_term_average")

var2.2 <- c("direction")

train.var1.2 <- TrainingSet[var1.2]
test.var1.2 <- ValidationSet[var1.2]
test.var1.2 <- test.var1.2[var1.2]
train.dep1.2 <- TrainingSet[var2.2]
test.dep1.2 <- ValidationSet[var2.2]

KNN.Multi.2 <- rep(NA,50)
for (i in 1:50){
  set.seed(1)
  inf.knn.pred <- knn(train.var1.2,test.var1.2,train.dep1.2$direction,k = i)
  KNN.Multi.2[i] <- mean(inf.knn.pred==test.dep1.2$direction)
}
KN2 <- c(1:50)
KNN.Multi.KN2 <- cbind(KNN.Multi.2,KN2)
inf.knn2 <- KNN.Multi.KN2[which.max(KNN.Multi.2), ]
inf.knn2
inf.knn.NoLag2 <- round(mean(inf.knn2[1])*100, 4)
inf.knn.NoLag2

#Success rate of 51.7997% when K = 32
```

```{r}
# Model 3 for Infation Indexed Long Term Average Direction
# LDA for Inflaction Indexed Long Term Average Direction

inf.d.lda.fit3 <- lda(direction~
                        prime_rate+
                        week_4_treasury_bill+
                        month_3_treasury_bill+
                        month_6_treasury_bill+
                        year_1_treasury_bill+
                        month_1_treasury_constant_maturity+
                        month_3_treasury_constant_maturity+
                        month_6_treasury_constant_maturity+
                        year_1_treasury_constant_maturity+
                        year_2_treasury_constant_maturity+
                        year_3_treasury_constant_maturity+
                        year_5_treasury_constant_maturity+
                        year_7_treasury_constant_maturity+
                        year_10_treasury_constant_maturity+
                        year_20_treasury_constant_maturity+
                        year_30_treasury_constant_maturity+
                        year_5_inflation_indexed_treasury_constant_maturity+
                        year_7_inflation_indexed_treasury_constant_maturity+
                        year_10_inflation_indexed_treasury_constant_maturity+
                        year_20_inflation_indexed_treasury_constant_maturity+
                        year_30_inflation_indexed_treasury_constant_maturity+
                        inflation_indexed_long_term_average, data=TrainingSet)

inf.d.lda.pred3 <- predict(inf.d.lda.fit3,ValidationSet)
inf.d.lda.class3 <- inf.d.lda.pred3$class
table(inf.d.lda.class3,inf.d.Val)
inf.d.LDA3 <- round(mean(inf.d.lda.class3==inf.d.Val)*100, 4)
inf.d.LDA3
#Success rate of 51.0172%

#QDA for Inflaction Indexed Long Term Average Direction
inf.d.qda.fit3 <- qda(direction~
                        prime_rate+
                        week_4_treasury_bill+
                        month_3_treasury_bill+
                        month_6_treasury_bill+
                        year_1_treasury_bill+
                        month_1_treasury_constant_maturity+
                        month_3_treasury_constant_maturity+
                        month_6_treasury_constant_maturity+
                        year_1_treasury_constant_maturity+
                        year_2_treasury_constant_maturity+
                        year_3_treasury_constant_maturity+
                        year_5_treasury_constant_maturity+
                        year_7_treasury_constant_maturity+
                        year_10_treasury_constant_maturity+
                        year_20_treasury_constant_maturity+
                        year_30_treasury_constant_maturity+
                        year_5_inflation_indexed_treasury_constant_maturity+
                        year_7_inflation_indexed_treasury_constant_maturity+
                        year_10_inflation_indexed_treasury_constant_maturity+
                        year_20_inflation_indexed_treasury_constant_maturity+
                        year_30_inflation_indexed_treasury_constant_maturity+
                        inflation_indexed_long_term_average, data=TrainingSet)

inf.d.qda.pred3 <- predict(inf.d.qda.fit3,ValidationSet)
inf.d.qda.class3 <- inf.d.qda.pred3$class
table(inf.d.qda.class3,inf.d.Val)
inf.d.QDA3 = round(mean(inf.d.qda.class3==inf.d.Val)*100, 4)
inf.d.QDA3
#Success rate of 50.3912%

#KNN for Inflaction Indexed Long Term Average Direction

var1.3 <- c("prime_rate",
             "week_4_treasury_bill",
             "month_3_treasury_bill",
             "month_6_treasury_bill",
             "year_1_treasury_bill",
             "month_1_treasury_constant_maturity" ,
             "month_3_treasury_constant_maturity",
             "month_6_treasury_constant_maturity" ,
             "year_1_treasury_constant_maturity",
             "year_2_treasury_constant_maturity",
             "year_3_treasury_constant_maturity",
             "year_5_treasury_constant_maturity",
             "year_7_treasury_constant_maturity",
             "year_10_treasury_constant_maturity",
             "year_20_treasury_constant_maturity",
             "year_30_treasury_constant_maturity",
             "year_5_inflation_indexed_treasury_constant_maturity", 
             "year_7_inflation_indexed_treasury_constant_maturity",
             "year_10_inflation_indexed_treasury_constant_maturity",
             "year_20_inflation_indexed_treasury_constant_maturity", 
             "year_30_inflation_indexed_treasury_constant_maturity",
             "inflation_indexed_long_term_average")

var2.3 <- c("direction")
train.var1.3 <- TrainingSet[var1.3]
test.var1.3 <- ValidationSet[var1.3]
test.var1.3 <- test.var1.3[var1.3]
train.dep1.3 <- TrainingSet[var2.3]
test.dep1.3 <- ValidationSet[var2.3]

KNN.Multi.3 <- rep(NA,50)
for (i in 1:50){
  set.seed(1)
  inf.knn.pred <- knn(train.var1.3,test.var1.3,train.dep1.3$direction,k = i)
  KNN.Multi.3[i] <- mean(inf.knn.pred==test.dep1.3$direction)
}
KN3 <- c(1:50)
KNN.Multi.KN3 <- cbind(KNN.Multi.3,KN3)
inf.knn3 <- KNN.Multi.KN3[which.max(KNN.Multi.3), ]
inf.knn3
inf.knn.NoLag3 <- round(mean(inf.knn3[1])*100, 4)
inf.knn.NoLag3

#Success rate of 51.4867% when K = 15
```

```{r}
# Model 4 for Infation Indexed Long Term Average Direction
# LDA for Inflaction Indexed Long Term Average Direction

inf.d.lda.fit4 <- lda(direction~
                        prime_rate+
                        week_4_treasury_bill+
                        month_3_treasury_bill+
                        month_6_treasury_bill+
                        year_1_treasury_bill+
                        month_1_treasury_constant_maturity+
                        month_3_treasury_constant_maturity+
                        month_6_treasury_constant_maturity+
                        year_1_treasury_constant_maturity+
                        year_2_treasury_constant_maturity+
                        year_3_treasury_constant_maturity+
                        year_5_treasury_constant_maturity+
                        year_7_treasury_constant_maturity+
                        year_10_treasury_constant_maturity+
                        year_20_treasury_constant_maturity+
                        year_30_treasury_constant_maturity, data=TrainingSet)

inf.d.lda.pred4 <- predict(inf.d.lda.fit4,ValidationSet)
inf.d.lda.class4 <- inf.d.lda.pred4$class
table(inf.d.lda.class4,inf.d.Val)
inf.d.LDA4 <- round(mean(inf.d.lda.class4==inf.d.Val)*100, 4)
inf.d.LDA4
#Success rate of 46.4789%

#QDA for Inflaction Indexed Long Term Average Direction
inf.d.qda.fit4 <- qda(direction~
                        prime_rate+
                        week_4_treasury_bill+
                        month_3_treasury_bill+
                        month_6_treasury_bill+
                        year_1_treasury_bill+
                        month_1_treasury_constant_maturity+
                        month_3_treasury_constant_maturity+
                        month_6_treasury_constant_maturity+
                        year_1_treasury_constant_maturity+
                        year_2_treasury_constant_maturity+
                        year_3_treasury_constant_maturity+
                        year_5_treasury_constant_maturity+
                        year_7_treasury_constant_maturity+
                        year_10_treasury_constant_maturity+
                        year_20_treasury_constant_maturity+
                        year_30_treasury_constant_maturity, data=TrainingSet)

inf.d.qda.pred4 <- predict(inf.d.qda.fit4,ValidationSet)
inf.d.qda.class4 <- inf.d.qda.pred4$class
table(inf.d.qda.class4,inf.d.Val)
inf.d.QDA4 = round(mean(inf.d.qda.class4==inf.d.Val)*100, 4)
inf.d.QDA4
#Success rate of 48.3568%

#KNN for Inflaction Indexed Long Term Average Direction

var1.4 <- c("prime_rate",
            "week_4_treasury_bill",
            "month_3_treasury_bill",
            "month_6_treasury_bill",
            "year_1_treasury_bill",
            "month_1_treasury_constant_maturity" ,
            "month_3_treasury_constant_maturity",
            "month_6_treasury_constant_maturity" ,
            "year_1_treasury_constant_maturity",
            "year_2_treasury_constant_maturity",
            "year_3_treasury_constant_maturity",
            "year_5_treasury_constant_maturity",
            "year_7_treasury_constant_maturity",
            "year_10_treasury_constant_maturity",
            "year_20_treasury_constant_maturity",
            "year_30_treasury_constant_maturity")

var2.4 <- c("direction")
train.var1.4 <- TrainingSet[var1.4]
test.var1.4 <- ValidationSet[var1.4]
test.var1.4 <- test.var1.4[var1.4]
train.dep1.4 <- TrainingSet[var2.4]
test.dep1.4 <- ValidationSet[var2.4]

KNN.Multi.4 <- rep(NA,50)
for (i in 1:50){
  set.seed(1)
  inf.knn.pred <- knn(train.var1.4,test.var1.4,train.dep1.4$direction,k = i)
  KNN.Multi.4[i] <- mean(inf.knn.pred==test.dep1.4$direction)
}
KN4 <- c(1:50)
KNN.Multi.KN4 <- cbind(KNN.Multi.4,KN4)
inf.knn4 <- KNN.Multi.KN4[which.max(KNN.Multi.4), ]
inf.knn4
inf.knn.NoLag4 <- round(mean(inf.knn4[1])*100, 4)
inf.knn.NoLag4

#Success rate of 49.9218% when K = 18
```

It appears that model 3 performs the best overall when running the LDA, QDA, and KNN on all of the models. It has the highest average percentage across the 4 models based on the training and validation that was run. 

#Question 4:

(Based on ISLR Chapter 9 #7) In this problem, you will use support vector approaches in order to
predict whether a given car gets high or low gas mileage based on the
Auto
data set.

##(a)
Create a binary variable that takes on a 1 for cars with gas mileage above the median, and a 0 for cars with gas mileage below the median.

```{r}
library(ISLR)
gsmed <- median(Auto$mpg)
Auto$mpglevel <- as.factor(ifelse(Auto$mpg > gsmed, 1, 0))
Auto$mpglevel 
```

##(b)
Fit a support vector classifier to the data with various values of cost, in order to predict whether a car gets high or low gas mileage. Report the cross-validation errors associated with different values of this parameter. Comment on your results.
```{r}

library(e1071)

set.seed(333)
tuning <- tune(svm, mpglevel ~ ., data = Auto, kernel = "linear", ranges = list(cost = c(0.01, 
    0.1, 1, 5, 10, 100)))
summary(tuning)
```
##(c)
Now repeat for (b), this time using SVMs with radial and polynomial basis kernels, with different values of gamma and degree and cost. Comment on your results.

```{r}

set.seed(333)
tuning1 <- tune(svm, mpglevel ~ ., data = Auto, kernel = "polynomial", ranges = list(cost = c(0.1, 
    1, 5, 10), degree = c(2, 3, 4)))
summary(tuning1)

set.seed(463)
tuning2 <- tune(svm, mpglevel ~ ., data = Auto, kernel = "radial", ranges = list(cost = c(0.1, 
    1, 5, 10), gamma = c(0.01, 0.1, 1, 5, 10, 100)))
summary(tuning2)
```

##(d)
Make some plots to back up your assertions in (b) and (c). Hint: In the lab, we used the plot() function for svm objects only in cases with p=2 When p>2,you can use the plot() function to create plots displaying pairs of variables at a time. Essentially, instead of typing plot(svmfit , dat) where svmfit contains your fitted model and dat is a data frame containing your data, you can type plot(svmfit , dat, x1???x4) in order to plot just the first and fourth variables. However, you must replace x1 and x4 with the correct variable names. To find out more, type ?plot.svm.

```{r}
svmrad <- svm(mpglevel ~ ., data = Auto, kernel = "radial", cost = 10, gamma = 0.01)
svmlin <- svm(mpglevel ~ ., data = Auto, kernel = "linear", cost = 1)
svmpol <- svm(mpglevel ~ ., data = Auto, kernel = "polynomial", cost = 10, degree = 2)

plotpairs = function(fit) {
    for (name in names(Auto)[!(names(Auto) %in% c("mpg", "mpglevel", "name"))]) {
        plot(fit, Auto, as.formula(paste("mpg~", name, sep = "")))
    }
}
```

```{r}
plotpairs(svmrad)
plotpairs(svmlin)
plotpairs(svmpol)

```

